---
title: "Practical Machine Learning"
output:
  html_document: default
  word_document: default
---

##Introduction:
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

The data for this project comes from this source: http://groupware.les.inf.puc-rio.br/har. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading in data
```{r}
raw_training <- read.csv('./pml-training.csv', header=T)
raw_testing <- read.csv('./pml-testing.csv', header=T)
```

Loading the caret packages
```{r}
library(caret)
```

Previewing the data to see if any columns are empty or can be excluded

str(raw_training)

I need to remove the columns that are empty or have NAs. I will not be including them in the model.
```{r}
cleanup <- sapply(names(raw_testing), function(x) all(is.na(raw_testing[,x])==TRUE))
nonames <- names(cleanup)[cleanup==FALSE]
nonames <- nonames [-(1:7)]
nonames <- nonames [1:(length(nonames)-1)]
```
##Cross Validation
Now I will separate the training data into 2 sets for cross validation.One set "70%" is used for the purpose of training and building the Random Forest model, and the other "30%" will be used to evaluate the model. 

I am Cross Validating the model by splitting the data into 2 specific sets.I split the data into 2 sets and will build the model on the larger set, and evaluate the model on both the large set and smaller set using 4 folds.
```{r}
training_sample <- createDataPartition(y=raw_training$classe, p=0.7, list=FALSE)
training <- raw_training[training_sample, ]
testing <- raw_training[-training_sample, ]
```

```{r}
fitControl <- trainControl(method='cv', number = 4)
```
Now I will build the model using the Random Forest model because I am the familiar with the model
``` {r}
rfmodel <- train(
                 classe ~ ., 
                 data=training[, c('classe', nonames)],
                 trControl=fitControl,
                 method='rf',
                 ntree=200)
rfmodel
```
As you can see, the Random Forest Model has a high accuracy so I think this model will be sufficient to use for the prediction of the 20 testing variables.



Now I will use a Confusion Matrix to describe the model against the Testing set and the Training Set
```{r}                 
prediction <- predict(rfmodel, newdata=testing)
ConfusionMatrix <- confusionMatrix(prediction, testing$classe)
print(ConfusionMatrix)

prediction <- predict(rfmodel, newdata=training)
ConfusionMatrix <- confusionMatrix(prediction, training$classe)
print(ConfusionMatrix)
```

##Out of Sample Error Rate
The out of sample error rate is very low. It is approximately 0 for the training set, and 0.92% for the subset testing set. 
Ex:  1 - .9908 = .0092

I want to build a decision tree model, even though the Random Forest model was strong.
```{r}
model_tree <- train(
   classe ~ ., 
   data=training[, c('classe', nonames)],
   trControl=fitControl,
   method='rpart')
 
 model_tree

prediction <- predict(model_tree, newdata=training)
ConfusionMatrix <- confusionMatrix(prediction, training$classe)
print(ConfusionMatrix)

```
This model only has a 0.49 accuracy  and a high out of sample error rate compared to the RF model so I will use the RF Model to predict the 20 observations


Finally, I will predict the classe of the 20 test observations in the raw testing data by using the model created.
```{r}
predictionTesting <- predict(rfmodel, newdata=raw_testing)
print(predictionTesting)
```
##Conclusion
I will use the results above to answer to 20 multiple choice questions. Based on the accuracy rates of the RF model, I think my results are fairly accurate to predict the sample observations.
After taking the multiple choice quiz, I answered all of the questions correctly.

